\chapter{Preface}
Most monographs and academic papers present constitutive models analytically with little to no emphasis on implementation, leaving readers to work out the details on their own.
However, implementing various models correctly, especially complex ones, can be challenging, since it requires not only a good understanding of their theoretical formulation but also a good grasp of the relevant numerical methods and programming details.

This book aims to provide a practical guide to the numerical implementation of plasticity models.

Readers are expected to have an undergraduate-level background in linear algebra, numerical analysis, elasticity, and some programming knowledge.
\section{Overview}
The book presents a collection of constitutive models, covering both uniaxial and triaxial versions for a wide range of common materials used in (civil/mechanical) engineering.
Implementation details are derived for each model.

The ultimate goal is to provide an easy-to-follow, error-free reference for readers who are interested in implementing constitutive models within modern conventional and novel plasticity frameworks.

A wide range of constitutive models (both formulations and implementations) are covered in this book, and some of them are very lengthy.
Accordingly, there are a huge number of symbols and formulae.

As the title indicates, this book is drafted as a reference book on constitutive modelling, and it should be self-contained in terms of theory.
For practical implementation, algorithms in pseudo code are given for most models.
Core \texttt{C++} implementations are also provided as references/comparisons for readers who are familiar with \texttt{C++} and scientific computation tools.
The code snippets are taken from the implementations in \texttt{suanPan} \cite{Chang2022}.
This book can also be used as a program manual.
\section{What Knowledge Is Needed?}
Constitutive modelling effectively concerns the numerical implementation of various plasticity theories.
Readers are expected to have a good undergraduate-level understanding of linear algebra and should feel comfortable with matrix/vector operations and analysis tools such as eigenanalysis.
Readers should also be fairly familiar with elasticity theories, especially small strain theory.
Furthermore, as can be seen later, plasticity theories can essentially be interpreted as differential equations; thus, a good grasp of numerical methods for differential equations (ODEs and PDEs) is also required.
However, it should be pointed out that, in most cases, only low order (very basic) methods are needed.

If 2D and 3D models are of interest, it is likely that a basic understanding of tensor algebra (literally I mean the first one or two chapters of any monographs on tensor algebra and nothing more) is also required.
Chapter \ref{chap:tensor} provides a brief summary of all tensor algebra knowledge needed in this book.
If one is able to understand the contents of Chapter \ref{chap:tensor}, that is sufficient.
\section{Implementation}
To get hands on with the actual implementation, one should then consider practical programming issues such as which programming language to use, which paradigm to follow, etc.
\subsection{Which Programming Language?}
In the early days, \texttt{C} and \texttt{Fortran} were the most popular languages for scientific computation.
One can easily find a large number of libraries written in either \texttt{C} or \texttt{Fortran} that are quietly driving the whole scientific computation community.

However, procedural programming may not necessarily be the ideal paradigm for scientific computation.
Consider, for example, that in those languages, vectors/matrices/tensors are represented by plain arrays.
The corresponding metadata, such as dimensions, can at best be stored in \texttt{struct}s.
Operating on arrays is cumbersome and error--prone.
As there is no operator overloading, all linear algebra operations have to be implemented as functions.
The implementation quickly becomes lengthy and hard to read, especially for 2D/3D elements and material models.

For \texttt{Fortran}, since matrix operations are natively supported, the corresponding implementation can be more concise.
However, when interfacing with external libraries (which is often necessary because they typically offer optimised performance), following the required signatures is often lengthy and tedious.
Furthermore, the language itself does not have a rich feature set. Writing \texttt{Fortran} is boring (my very subjective opinion) and often ends up being one dull loop after another.

When choosing a proper language for scientific computation, the following three aspects are important.
\begin{itemize}
\item Operator Overloading\\
Operator overloading is a mechanism that allows one to define custom behaviour for standard operators (e.g., $+$, $-$, $*$, $/$.) when applied to custom types.
This is particularly useful in scientific computing, where mathematical operations on vectors, matrices, and tensors are common.
With operator overloading, one can write code that more closely resembles mathematical notation, making it more readable and easier to understand.
It may also offer performance optimisation.
Consider the multiplication of three matrices; with operator overloading, it can be concisely expressed as
\begin{cppcode}
    mat A, B, C;
    auto D = A * B * C;
\end{cppcode}
However, if the chosen language does not support it, this can only be done via conventional function calls.
\begin{cppcode}
    mat matmul(const mat&, const mat&);
    mat A, B, C;
    auto D = matmul(matmul(A, B), C);
\end{cppcode}
\item Ecosystem\\
For implementing standalone material models, whether there is a rich ecosystem may not be a big concern.
However, at higher levels, material models need to be integrated into larger analysis frameworks, and a rich ecosystem can significantly reduce reinvention and the overall workload.
\item Multiparadigm Support\\
Although not strictly necessary, multiparadigm support can be very useful in scientific computation.
For finite element analysis, nodes/elements can be naturally modelled as objects, thus object--oriented programming (OOP) is a natural choice.
\end{itemize}

In this book, \texttt{C++} is chosen as the implementation language.
It may not be the perfect choice for everyone (subject to personal preference), but it is definitely a reasonable one.
A similar mindset is also adopted by other authors \cite{Poeschel2010}.
Nevertheless, one is free to implement the presented algorithms in any programming language of their choice.
\subsection{Notes on Miscellaneous Things}
\subsubsection{State Determination}
In all the state determination algorithms presented in this book, it is assumed by default that all local iterations eventually converge.
However, this is often not what happens in practice.
For example, most algorithms solve a local nonlinear system with the Newton--Raphson method, which is known to only have a local convergence guarantee.
This implies that the algorithm would simply fail if the initial starting point is far enough from the desired solution.
To improve robustness, it is always possible to add more numerical procedures to safeguard the algorithm.
Thus, the presented implementations are the `reference' ones, not the `golden' ones.
\subsubsection{Internal Storage}
Mathematical objects, such as tensors, can be faithfully represented in programs just as they are in mathematics.
This means, for example, that second order tensors can be represented as 2D matrices, and fourth order tensors can be represented as 4D cubes.
A complete set of rules can then be defined to perform tensor algebra.

This can be seen in many other applications.
However, under the Cauchy stress theory and with the Voigt notation, row/column vectors and 2D matrices are sufficient to implement most models.

In this book, all models are implemented in this way.