\section{Orthotropic Hoffman Model}\label{sec:hoffman}
Here we introduce an anisotropic model that adopts the orthotropic Hoffman yielding criterion.
This framework resembles the isotropic von Mises model.
In fact, the von Mises model can be seen as a degenerate version of the Hoffman model.
It can be used to model orthotropic materials such as sheet steel and timber.
See also the corresponding section (\S~10.3) in \cite{SouzaNeto2008}.
\subsection{Theory}
\subsubsection{Yield Function}
The yield function adopts the Hoffman criterion, which is expressed in compressed matrix form,
\begin{gather}\label{eq:orthotropic_yield_function_proto}
    f=\dfrac{1}{2}\bsigma^\mT\mb{P}\bsigma+\mb{q}^\mT\bsigma-\sigma_y^2,
\end{gather}
where $\mb{P}=\mb{P}^\mT$ and $\bq$ denote a constant scaling matrix and vector.
Various other orthotropic yielding criteria can also fit this form; see \cite{Oller2003} for details.
Apparently, $\mb{P}$ and $\bq$ represent two tensors of order four and two, respectively.
But since \cite{Oller2003} uses matrix notation, we follow that convention.
For example, the Hoffman criterion can be expressed as
\begin{tiny}
    \begin{gather}\label{eq:hoffman_criterion}
        \dfrac{1}{2}\mb{P}=\begin{bmatrix}
            T_1                     & -\dfrac{T_1+T_2-T_3}{2} & -\dfrac{T_3+T_1-T_2}{2} &                     &                     &                     \\[4mm]
            -\dfrac{T_1+T_2-T_3}{2} & T_2                     & -\dfrac{T_2+T_3-T_1}{2} &                     &                     &                     \\[4mm]
            -\dfrac{T_3+T_1-T_2}{2} & -\dfrac{T_2+T_3-T_1}{2} & T_3                     &                     &                     &                     \\[4mm]
                                    &                         &                         & \dfrac{1}{f_{12}^2} &                     &                     \\[4mm]
                                    &                         &                         &                     & \dfrac{1}{f_{23}^2} &                     \\[4mm]
                                    &                         &                         &                     &                     & \dfrac{1}{f_{13}^2}
        \end{bmatrix},\quad
        \mb{q}=\begin{bmatrix}
            \left(f_{11}^c-f_{11}^t\right)T_1 \\[4mm]
            \left(f_{22}^c-f_{22}^t\right)T_2 \\[4mm]
            \left(f_{33}^c-f_{33}^t\right)T_3 \\[4mm]
            0                                 \\[4mm]
            0                                 \\[4mm]
            0
        \end{bmatrix},
    \end{gather}
\end{tiny}
in which,
\begin{gather}
    T_1=\dfrac{1}{f_{11}^tf_{11}^c},\qquad
    T_2=\dfrac{1}{f_{22}^tf_{22}^c},\qquad
    T_3=\dfrac{1}{f_{33}^tf_{33}^c},
\end{gather}
with $f_{ij}^\aleph$ representing the yielding stress along different directions.

Unlike the other models introduced in this book, the yield function here is expressed in matrix form, so the following discussion is mainly based on matrix notation.
An equivalent tensor notation may also exist, but since it is not used in the literature, we do not introduce it here.
\subsubsection{Flow Rule}
Associated plasticity is assumed, so the plastic potential $g$ is identical to $f$. The plastic flow direction is then
\begin{gather}
    \bn=\pdfrac{g}{\bsigma}=\pdfrac{f}{\bsigma}=\mb{P}\bsigma+\mb{q}.
\end{gather}
The flow rule can be defined as
\begin{gather}
    \dot{\bvarepsilon^p}=\gamma\bn.
\end{gather}
It is worth noting that since the yield function is expressed as a function of the stress (Voigt) vector rather than the stress tensor, the partial derivative $\mb{P}\bsigma+\mb{q}$ is the derivative with regard to the vector $\bsigma$, which already contains the scaling factor for shear terms.
This is different from other models that use the tensor notation.
\subsubsection{Hardening Law}
The reference yield stress $\sigma_y$ is defined as a function of the accumulated equivalent plastic strain $\varepsilon_{p}$.
\begin{gather}
    \sigma_y=H\left(\varepsilon^{p}\right).
\end{gather}
It is called the `reference' yield stress here --- there is no unique uniaxial equivalence because the yield stress may vary with direction.

The evolution of $\varepsilon^{p}$ is defined by
\begin{gather}
    \dot{\varepsilon^p}=\sqrt{\dfrac{2}{3}\dot{\bvarepsilon^p}:\dot{\bvarepsilon^p}}=\gamma\sqrt{\dfrac{2}{3}\bn:\bn},
\end{gather}
where $\bn:\bn$ should be treated as the double contraction of strain tensors, that is,
\begin{gather}
    \underbrace{\bn:\bn}_\text{tensor}=\underbrace{\bn^\mT\mb{T}\bn}_\text{vector},
\end{gather}
with $\mb{T}=\diag{\begin{matrix}
            1 & 1 & 1 & \dfrac{1}{2} & \dfrac{1}{2} & \dfrac{1}{2}
        \end{matrix}}$.
This is due to the fact that \eqsref{eq:orthotropic_yield_function_proto} is expressed in vectors/matrices rather than tensors!

Although \textbf{not} correct, we'll use $\norm{\bn}_e$ to represent $\sqrt{\dfrac{2}{3}\bn:\bn}$ for brevity in the following.
\subsection{Formulation}
To some extent, the model is even simpler than the von Mises model as there is no back stress to support kinematic hardening.
This is reasonable as there is no experimental evidence that timber exhibits the Bauschinger effect.
For other materials, specific models can be developed, in which kinematic hardening can be introduced.
Furthermore, the yield function involves only matrix--vector operations, so the corresponding derivatives are relatively easy to compute.
\begin{table}[ht]
    \centering
    \begin{tabular}{rl}
        \toprule
        Constitutive Law & $\bsigma=\mb{E}\left(\bvarepsilon-\bvarepsilon^p\right)$              \\
        Yield Function   & $f=\dfrac{1}{2}\bsigma^\mT\mb{P}\bsigma+\mb{q}^\mT\bsigma-\sigma_y^2$ \\
        Flow Rule        & $\dot{\bvarepsilon^p}=\gamma\bn$                                      \\
        Hardening Law    & $\dot{\varepsilon^p}=\gamma\norm{\bn}_e$                              \\\bottomrule
    \end{tabular}
\end{table}
\subsubsection{Elastic Loading/Unloading}
Compared with the von Mises framework, there is no essential difference in elastic loading/unloading. Plasticity is frozen at the beginning of each substep, which allows one to compute the trial stress as
\begin{gather}
    \bsigma^\text{trial}=\mb{E}\left(\bvarepsilon_{n+1}-\bvarepsilon^p_n\right)\xcancel{=\bsigma_n+\mb{E}\left(\bvarepsilon_{n+1}-\bvarepsilon_n\right)}.
\end{gather}
Note that the second form is not used here.
In most cases, both forms are equivalent.
However, some models may further apply damage mechanics to the result of plasticity, making the second form incorrect (as $\bsigma_n$ may contain damage reductions).
Here we demonstrate the usage of the first form.

The trial yield function can then be computed using the unchanged plastic strain.
\begin{gather}
    f^\text{trial}=\dfrac{1}{2}\bsigma^\text{trial,T}\mb{P}\bsigma^\text{trial}+\mb{q}^\mT\bsigma^\text{trial}-\sigma_{y,n}^2.
\end{gather}
\subsubsection{Plastic Evolution}
The discretised evolution of plastic strain is written as
\begin{gather}
    \bvarepsilon^p_{n+1}=\bvarepsilon^p_{n}+\Delta\bvarepsilon^p=\bvarepsilon^p_{n}+\gamma\bn_{m}.
\end{gather}
Note that we do not directly use $\bn_{n+1}$ here. Instead, we use $\bn_{m}$ to allow multiple implementations.
Specifically, if $\bn_{m}=\bn_{n+1}$, it is effectively the implicit Euler method.
\subsubsection{Local Residual}
The residual is chosen as follows. For brevity, all subscripts $\left(\cdot\right)_{n+1}$ are dropped.
\begin{gather}
    \mb{R}=\left\{
    \begin{array}{l}
        \dfrac{1}{2}\bsigma^\mT\mb{P}\bsigma+\mb{q}^\mT\bsigma-\sigma_y^2, \\[4mm]
        \bsigma+\gamma\mb{E}\bn_m-\bsigma^\text{trial}.
    \end{array}
    \right.
\end{gather}
We use $\mb{E}$ to denote the elasticity tensor $\mb{D}$ expressed in matrix form.
They represent the same quantity.

By choosing $\mb{x}=\begin{bmatrix}
        \gamma & \bsigma
    \end{bmatrix}^\mT$ as the independent variables, the Jacobian can then be computed as
\begin{gather}
    \mb{J}=\pdfrac{\mb{R}}{\mb{x}}=\begin{bmatrix}
        -2\sigma_y\ddfrac{\sigma_y}{\varepsilon_{p}}\norm{\bn_m}_e & \bn^\mT-2\sigma_y\ddfrac{\sigma_y}{\varepsilon_{p}}\gamma\ddfrac{\norm{\bn_m}_e}{\bsigma} \\[4mm]
        \mb{E}\bn_m                                                & \mb{I}+\gamma\mb{E}\ddfrac{\bn_m}{\bsigma}
    \end{bmatrix}.
\end{gather}

Some references further derive a scalar local residual at the cost of a more complicated gradient. Here we choose to increase the size of the local system in order to express the Jacobian in a simpler form. Performance-wise, a scalar local residual does not necessarily lead to faster state determination.
\subsubsection{Consistent Tangent Stiffness}
The consistent tangent stiffness can be computed directly from the local residual, given that $\bsigma_{n+1}$ is chosen as the variable. Differentiating $\mb{R}$ at equilibrium/convergence, where $\mb{R}=\mb{0}$, gives
\begin{gather}\label{eq:universal_consistent_tangent}
    \pdfrac{\mb{R}}{\mb{x}}\pdfrac{\mb{x}}{\bvarepsilon_{n+1}}+\pdfrac{\mb{R}}{\bvarepsilon_{n+1}}=\mb{0},
\end{gather}
which, after rearrangement, gives
\begin{gather}
    \pdfrac{\mb{x}}{\bvarepsilon_{n+1}}=-\left(\pdfrac{\mb{R}}{\mb{x}}\right)^{-1}\pdfrac{\mb{R}}{\bvarepsilon_{n+1}}=-\mb{J}^{-1}\pdfrac{\mb{R}}{\bvarepsilon_{n+1}}.
\end{gather}

Since the left-hand side $\pdfrac{\mb{x}}{\bvarepsilon_{n+1}}$ contains
\begin{gather}
    \pdfrac{\mb{x}}{\bvarepsilon_{n+1}}=\begin{bmatrix}
        \pdfrac{\gamma}{\bvarepsilon_{n+1}} \\[4mm]
        \pdfrac{\bsigma_{n+1}}{\bvarepsilon_{n+1}}
    \end{bmatrix},
\end{gather}
then,
\begin{gather}\label{eq:ortho_consistent_tangent}
    \pdfrac{\bsigma_{n+1}}{\bvarepsilon_{n+1}}=-\left(\mb{J}^{-1}\pdfrac{\mb{R}}{\bvarepsilon_{n+1}}\right)^{\langle2-7\rangle},
\end{gather}
where $\left(\cdot\right)^{\langle2-7\rangle}$ denotes the second to the seventh rows of the target quantity $\left(\cdot\right)$.

Unlike the von Mises framework, where an analytical expression for the consistent tangent stiffness matrix is derived, here we take advantage of the fact that when local equilibrium is achieved, $\mb{R}=\mb{0}$ or at least $\mb{R}\approx\mb{0}$. This allows full differentiation to obtain useful quantities that may otherwise be difficult to compute. If $\bsigma_{n+1}$ is directly involved as one of the independent variables in the local iteration, the consistent stiffness can be obtained directly. Otherwise, additional simple steps (chain rule) often need to be applied to the stress update formula to compute $\pdfrac{\bsigma_{n+1}}{\bvarepsilon_{n+1}}$.

This method avoids the computation of lengthy, cumbersome analytical expressions for the consistent tangent. In most cases, it is also very simple to implement because $\mb{J}$ is already available when the local iteration converges, and $\pdfrac{\mb{R}}{\bvarepsilon_{n+1}}$ is often very easy to compute. Readers should try to grasp the beauty of \eqsref{eq:universal_consistent_tangent}, as this method will be used frequently in the models introduced later in this book.
\subsection{Implementation}
\subsubsection{Trapezoidal Integration}
In this particular model, we demonstrate the less common trapezoidal rule.
By the trapezoidal rule, we mean that the update of the plastic strain should be expressed as
\begin{gather}
    \bvarepsilon^p_{n+1}=\bvarepsilon^p_{n}+\gamma\bn_{m},
\end{gather}
as presented earlier.
In a conventional context, it can be further expressed as
\begin{gather}
    \bvarepsilon^p_{n+1}=\bvarepsilon^p_{n}+\gamma\dfrac{\bn_{n}+\bn_{n+1}}{2},\qquad\text{(WRONG!)}
\end{gather}
where $\bn_{n}$ and $\bn_{n+1}$ are the flow directions at the beginning and the end of the step, respectively.
However, one must notice that the current state $\left(\cdot\right)_n$ may be elastic!
The above expression is correct only when the current state is on the yield surface.
To be precise, as we know that the new state $\left(\cdot\right)_{n+1}$ must be plastic, the above expression has to be modified as
\begin{gather}
    \bvarepsilon^p_{n+1}=\bvarepsilon^p_{n}+\gamma\dfrac{\bn_\text{onset}+\bn_{n+1}}{2},
\end{gather}
where $\bn_\text{onset}$ is the flow direction at the onset of plasticity in the current step.
It thus implies that
\begin{gather}
    \bn_{m}=\dfrac{\bn_\text{onset}+\bn_{n+1}}{2}.
\end{gather}

Assuming proportional loading, finding $\bn_\text{onset}$ is equivalent to finding a ratio $r$ such that the onset stress $\bsigma_\text{onset}$, which is defined as
\begin{gather}
    \bsigma_\text{onset}=\mb{E}\left(\bvarepsilon_n+r\Delta\bvarepsilon-\bvarepsilon_n^p\right),
\end{gather}
is on the yield surface, that is,
\begin{gather}
    f_\text{onset}=f\left(\bsigma_\text{onset}\right)=\dfrac{1}{2}\bsigma_\text{onset}^\mT\mb{P}\bsigma_\text{onset}+\mb{q}^\mT\bsigma_\text{onset}-\sigma_{y,n}^2=0.
\end{gather}
Note that in this procedure plasticity is frozen; the only variable is the ratio $r$.
This is a well-bounded problem --- $r$ is guaranteed to be greater than zero and less than one.
The solution $r$ can be computed using a root-finding method.
Alternatively, since $f_\text{onset}$ is a quadratic function of $r$, one can compute the roots of the equation directly.
Once $r$ is known, $\bsigma_\text{onset}$ can then be computed, as well as $\bn_\text{onset}$.

It is also necessary to compute the derivative.
From the equilibrium of $f_\text{onset}$, one has
\begin{gather}
    \pdfrac{f_\text{onset}}{r}\md{r}+\pdfrac{f_\text{onset}}{\bvarepsilon_{n+1}}\md{\bvarepsilon_{n+1}}=0,
\end{gather}
thus,
\begin{gather}
    \ddfrac{r}{\bvarepsilon_{n+1}}=\dfrac{-1}{\pdfrac{f_\text{onset}}{r}}\pdfrac{f_\text{onset}}{\bvarepsilon_{n+1}},
\end{gather}
with
\begin{gather}
    \pdfrac{f_\text{onset}}{r}=\bn_\text{onset}^\mT\mb{E}\Delta\bvarepsilon,\qquad
    \pdfrac{f_\text{onset}}{\bvarepsilon_{n+1}}=\bn_\text{onset}^\mT\mb{E}r.
\end{gather}
With the above, one can further derive
\begin{gather}
    \ddfrac{\bn_\text{onset}}{\bvarepsilon_{n+1}}=\mb{PE}\left(\Delta\bvarepsilon\ddfrac{r}{\bvarepsilon_{n+1}}+r\mb{I}\right)=r\left(\mb{PE}-\mb{PE}\dfrac{\Delta\bvarepsilon\bn_\text{onset}^\mT\mb{E}}{\bn_\text{onset}^\mT\mb{E}\Delta\bvarepsilon}\right).
\end{gather}
A slight rearrangement gives
\begin{gather}
    \ddfrac{\bn_\text{onset}}{\bvarepsilon_{n+1}}=r\mb{P}\left(\mb{I}-\dfrac{\mb{E}\Delta\bvarepsilon\bn_\text{onset}^\mT}{\bn_\text{onset}^\mT\mb{E}\Delta\bvarepsilon}\right)\mb{E}=r\mb{P}\left(\mb{I}-\dfrac{\Delta\bsigma\bn_\text{onset}^\mT}{\bn_\text{onset}^\mT\Delta\bsigma}\right)\mb{E}.
\end{gather}
It is clear that when $r\rightarrow0$, the derivative approaches zero as $\ddfrac{\bn_\text{onset}}{\bvarepsilon_{n+1}}\rightarrow\mb{0}$.

With $\bn_\text{onset}$, it is possible to compute the explicit expression of $\pdfrac{\mb{R}}{\bvarepsilon_{n+1}}$ as
\begin{gather}
    \pdfrac{\mb{R}}{\bvarepsilon_{n+1}}=\begin{bmatrix}
        -2\sigma_y\ddfrac{\sigma_y}{\varepsilon_{p}}\gamma\ddfrac{\norm{\bn_m}_e}{\bvarepsilon_{n+1}} \\[4mm]
        \gamma\mb{E}\ddfrac{\bn_m}{\bvarepsilon_{n+1}}-\mb{E}
    \end{bmatrix}=\begin{bmatrix}
        -\sigma_y\ddfrac{\sigma_y}{\varepsilon_{p}}\gamma\ddfrac{\norm{\bn_m}_e}{\bn_m}r\mb{P}\left(\mb{I}-\dfrac{\Delta\bsigma\bn_\text{onset}^\mT}{\bn_\text{onset}^\mT\Delta\bsigma}\right)\mb{E} \\[4mm]
        \dfrac{\gamma}{2}\mb{E}r\mb{P}\left(\mb{I}-\dfrac{\Delta\bsigma\bn_\text{onset}^\mT}{\bn_\text{onset}^\mT\Delta\bsigma}\right)\mb{E}-\mb{E}
    \end{bmatrix}.
\end{gather}
Rearranging the above gives
\begin{gather}
    \pdfrac{\mb{R}}{\bvarepsilon_{n+1}}=\begin{bmatrix}
        -\sigma_yr\gamma\ddfrac{\sigma_y}{\varepsilon_{p}}\ddfrac{\norm{\bn_m}_e}{\bn_m}\mb{P}\left(\mb{I}-\dfrac{\Delta\bsigma\bn_\text{onset}^\mT}{\bn_\text{onset}^\mT\Delta\bsigma}\right) \\[4mm]
        \dfrac{r\gamma}{2}\mb{EP}\left(\mb{I}-\dfrac{\Delta\bsigma\bn_\text{onset}^\mT}{\bn_\text{onset}^\mT\Delta\bsigma}\right)-\mb{I}
    \end{bmatrix}\mb{E}.
\end{gather}
The above can be inserted into \eqsref{eq:ortho_consistent_tangent} to obtain the consistent tangent stiffness.
There is no tensor algebra involved; all expressions are in matrix form.
\subsubsection{Line Search}
Since it is an anisotropic model, the local iteration may have difficulties in convergence, especially when a high anisotropy is defined.
Some implementations \cite{Krasnovskiy2004,SouzaNeto2008,Scherzinger2017} adopt line search, which effectively becomes a damped Newton--Raphson method.
It indeed improves numerical stability in certain cases, but does not guarantee global convergence.
The additional cost of searching is not necessarily negligible, as a desired step size is not guaranteed to be found in a fixed number of searches.
What we end up with is a more costly algorithm that may not converge.

Instead of line search, noting that there must exist a solution for some $\gamma>0$, it is possible to use a bracketing method that guarantees a solution can be found for arbitrarily specified accuracy.
The idea can be explained as follows.
For a given $\gamma$, it is possible to compute the stress $\bsigma$ via
\begin{gather}
    \bsigma=\bsigma^\text{trial}-\gamma\mb{E}\bn_m,
\end{gather}
thus
\begin{multline}
    \left(\mb{I}+\dfrac{\gamma}{2}\mb{E}\mb{P}\right)\bsigma=\bsigma^\text{trial}-\dfrac{\gamma}{2}\mb{E}\left(\bn_\text{onset}+\mb{q}\right) \\\longrightarrow\qquad
    \bsigma=\left(\mb{I}+\dfrac{\gamma}{2}\mb{E}\mb{P}\right)^{-1}\left(\bsigma^\text{trial}-\dfrac{\gamma}{2}\mb{E}\left(\bn_\text{onset}+\mb{q}\right)\right).
\end{multline}
Once $\bsigma$ is known, the yield function can be evaluated.
This effectively turns $f=f\left(\gamma\right)$ into a univariate function of $\gamma$. Using the above procedure, no iteration is required to evaluate $f$ for a given $\gamma$, and the evaluation is precise.
Knowing that $f\left(0\right)>0$, one may start with a small $\gamma_b>0$ and compute $f\left(\gamma_b\right)$ before applying a root-finding method.
If $f\left(\gamma_b\right)>0$, double $\gamma_b$ by, for example, $\gamma_b\leftarrow2\gamma_b$ and continue evaluating $f\left(\gamma_b\right)$ until $f\left(\gamma_b\right)\leqslant0$.
This ensures that the solution is bracketed by $0$ and $\gamma_b$.
One can then use root-finding methods to find the solution $\gamma$.
\subsubsection{Return Mapping Implementation}
The following is a reference implementation that switches between the implicit Euler method and the trapezoidal method based on the size of the strain increment.
The criterion used to determine which method to use is purely heuristic, and it is not guaranteed to be optimal.
The general idea is to use the second-order method when the strain increment is small so that asymptotic convergence can be retained.
\begin{cppcode}
int NonlinearOrthotropic::update_trial_status(const vec& t_strain) {
    incre_strain = (trial_strain = t_strain) - current_strain;

    const auto norm_incre_strain = tensor::strain::norm(incre_strain);

    if(norm_incre_strain <= datum::eps) return SUANPAN_SUCCESS;

    return norm_incre_strain < reference_strain ? trapezoidal_return() : euler_return();
}
\end{cppcode}

The two integration methods follow a similar structure.
Note the computation of the onset state.
\begin{cppcode}
int NonlinearOrthotropic::trapezoidal_return() {
    trial_history = current_history;
    const auto& current_ep = current_history(0);
    auto& ep = trial_history(0);
    vec plastic_strain(&trial_history(1), 6, false, true);

    const vec6 base_s = (trial_stiffness = initial_stiffness) * (current_strain - plastic_strain);
    const vec6 incre_s = initial_stiffness * incre_strain;
    const vec6 predictor_s = trial_stress = base_s + incre_s;

    const auto current_k = compute_k(current_ep);

    // elastic region
    if(ortho_inner(predictor_s) <= current_k * current_k) return SUANPAN_SUCCESS;

    vec6 onset_n = proj_p * base_s + proj_q;
    mat66 dnds;

    const auto base_f = ortho_inner(base_s) - current_k * current_k;
    const auto inside_surface = std::signbit(base_f);
    if(inside_surface) {
        // elastic loading to plastic
        // need to find the onset point
        const auto quadratic_a = dot(incre_s, proj_p * incre_s);
        const auto quadratic_b = dot(incre_s, onset_n);
        const auto onset_r = (std::sqrt(quadratic_b * quadratic_b - 2. * quadratic_a * base_f) - quadratic_b) / quadratic_a;
        onset_n += proj_p * incre_s * onset_r;
        dnds = proj_p * (onset_r * eye(6, 6) - onset_r / dot(onset_n, incre_s) * incre_s * onset_n.t());
    }

    auto gamma{0.};

    vec7 incre(fill::none), residual(fill::none);
    mat77 jacobian(fill::none);

    auto counter{0u};
    auto try_bisection{false};
    while(true) {
        if(max_iteration == ++counter) {
            try_bisection = true;

            const vec6 const_a = onset_n + proj_q;
            const vec6 const_b = initial_stiffness * const_a;

            const auto approx_update = [&](const double gm) {
                gamma = gm;
                const vec6 approx_a = proj_p * (trial_stress = solve(eye(6, 6) + .5 * gamma * elastic_p, predictor_s - .5 * gamma * const_b));
                const auto approx_k = compute_k(ep = current_ep + .5 * gamma * root_two_third * tensor::strain::norm(approx_a + const_a));
                return .5 * dot(trial_stress, approx_a) + dot(trial_stress, proj_q) - approx_k * approx_k;
            };

            const auto norm_incre_strain = tensor::strain::norm(incre_strain);

            ridders_guess(approx_update, 0., .25 * norm_incre_strain / tensor::strain::norm(proj_p * predictor_s + proj_q), tolerance);

            if(gamma * tensor::strain::norm(proj_p * trial_stress + const_a) > 2. * norm_incre_strain) {
                suanpan_error("Cannot converge within {} iterations.\n", max_iteration);
                return SUANPAN_FAIL;
            }
        }

        const vec6 n = proj_p * trial_stress + proj_q;
        const vec6 n_mid = .5 * (n + onset_n);
        const auto norm_n_mid = root_two_third * tensor::strain::norm(n_mid);
        const rowvec6 pnn = .5 * two_third / norm_n_mid * (n_mid % tensor::strain::norm_weight).t(); // part of some derivative
        const auto k = compute_k(ep = current_ep + gamma * norm_n_mid);

        residual(sa) = ortho_inner(trial_stress) - k * k;
        residual(sb) = trial_stress + gamma * initial_stiffness * n_mid - predictor_s;

        const auto pk = -2. * k * compute_dk(ep);

        jacobian(sa, sa) = pk * norm_n_mid;
        jacobian(sa, sb) = n.t() + pk * gamma * pnn * proj_p;
        jacobian(sb, sa) = initial_stiffness * n_mid;
        jacobian(sb, sb) = eye(6, 6) + .5 * gamma * elastic_p;

        if(!solve(incre, jacobian, residual, solve_opts::equilibrate)) return SUANPAN_FAIL;

        const auto error = suanpan::inf_norm(incre);
        suanpan_debug("Local plasticity iteration (2nd) error: {:.5E}.\n", error);

        if(error < tolerance * (1. + yield_stress.max()) || (suanpan::inf_norm(residual) < tolerance && counter > 5u) || try_bisection) {
            plastic_strain += gamma * n_mid;

            mat::fixed<7, 6> left(fill::none), right(fill::zeros);
            if(inside_surface) {
                right.row(sa) = -pk * gamma * pnn * dnds;
                right.rows(sb) = eye(6, 6) - .5 * gamma * initial_stiffness * dnds;
            }
            else right.rows(sb) = eye(6, 6);

            if(!solve(left, jacobian, right, solve_opts::equilibrate)) return SUANPAN_FAIL;

            trial_stiffness = left.rows(sb) * initial_stiffness;

            return SUANPAN_SUCCESS;
        }

        gamma = std::max(0., gamma - incre(sa));
        trial_stress -= incre(sb);
    }
}
\end{cppcode}

The Euler method is simpler.
\begin{cppcode}
int NonlinearOrthotropic::euler_return() {
    trial_history = current_history;
    const auto& current_ep = current_history(0);
    auto& ep = trial_history(0);
    vec plastic_strain(&trial_history(1), 6, false, true);

    const vec6 predictor_s = trial_stress = (trial_stiffness = initial_stiffness) * (trial_strain - plastic_strain);

    // elastic region
    if(ortho_inner(predictor_s) <= std::pow(compute_k(current_ep), 2.)) return SUANPAN_SUCCESS;

    auto gamma{0.};

    vec7 incre(fill::none), residual(fill::none);
    mat77 jacobian(fill::none);

    auto counter{0u};
    auto try_bisection{false};
    while(true) {
        if(max_iteration == ++counter) {
            try_bisection = true;

            const vec6 elastic_q = initial_stiffness * proj_q;

            const auto approx_update = [&](const double gm) {
                gamma = gm;
                const vec6 approx_a = proj_p * (trial_stress = solve(eye(6, 6) + gamma * elastic_p, predictor_s - gamma * elastic_q));
                const auto approx_k = compute_k(ep = current_ep + gamma * root_two_third * tensor::strain::norm(approx_a + proj_q));
                return .5 * dot(trial_stress, approx_a) + dot(trial_stress, proj_q) - approx_k * approx_k;
            };

            const auto norm_incre_strain = tensor::strain::norm(incre_strain);

            ridders_guess(approx_update, 0., .25 * norm_incre_strain / tensor::strain::norm(proj_p * predictor_s + proj_q), tolerance);

            if(gamma * tensor::strain::norm(proj_p * trial_stress + proj_q) > norm_incre_strain) {
                suanpan_error("Cannot converge within {} iterations.\n", max_iteration);
                return SUANPAN_FAIL;
            }
        }

        const vec6 n = proj_p * trial_stress + proj_q;
        const auto norm_n = root_two_third * tensor::strain::norm(n);
        const rowvec6 pnn = two_third / norm_n * (n % tensor::strain::norm_weight).t(); // part of some derivative
        const auto k = compute_k(ep = current_ep + gamma * norm_n);

        residual(sa) = ortho_inner(trial_stress) - k * k;
        residual(sb) = trial_stress + gamma * initial_stiffness * n - predictor_s;

        const auto pk = -2. * k * compute_dk(ep);

        jacobian(sa, sa) = pk * norm_n;
        jacobian(sa, sb) = n.t() + pk * gamma * pnn * proj_p;
        jacobian(sb, sa) = initial_stiffness * n;
        jacobian(sb, sb) = eye(6, 6) + gamma * elastic_p;

        if(!solve(incre, jacobian, residual, solve_opts::equilibrate)) return SUANPAN_FAIL;

        const auto error = suanpan::inf_norm(incre);
        suanpan_debug("Local plasticity iteration (1st) error: {:.5E}.\n", error);

        if(error < tolerance * (1. + yield_stress.max()) || (suanpan::inf_norm(residual) < tolerance && counter > 5u) || try_bisection) {
            plastic_strain += gamma * n;

            mat::fixed<7, 6> left(fill::none), right(fill::zeros);
            right.rows(sb) = initial_stiffness;

            if(!solve(left, jacobian, right, solve_opts::equilibrate)) return SUANPAN_FAIL;

            trial_stiffness = left.rows(sb);

            return SUANPAN_SUCCESS;
        }

        gamma = std::max(0., gamma - incre(sa));
        trial_stress -= incre(sb);
    }
}
\end{cppcode}
