\section{von Mises Framework}
Here, the uniaxial combined isotropic/kinematic model introduced in \secref{sec:isotropic/kinematic} is reformulated in 3D space. Some differences will appear, but the final local residual remains a scalar equation.
\subsection{Theory}
\subsubsection{Yield Function}
The von Mises yield criterion is adopted,
\begin{gather}
    f=\norm{\beeta}-\sqrt{\dfrac{2}{3}}\sigma^y,
\end{gather}
where $\beeta=\bs-\balpha$ is the shifted stress, $\balpha$ denotes the back stress, and $\bs=\dev{\bsigma}$ denotes the deviatoric stress.
By definition, the back stress $\balpha$ is also deviatoric, thus $\tr{\balpha}=0$.
It shifts the centre of the yield surface in the deviatoric stress space.

The only purpose of the constant $\sqrt{\dfrac{2}{3}}$ is to produce a consistent response under uniaxial loading with the same set of model parameters.
To see this, consider the uniaxial loading case, which can be represented by the following stress tensor
\begin{gather}
    \bsigma=\begin{bmatrix}
        \sigma & 0 & 0 \\
        0      & 0 & 0 \\
        0      & 0 & 0
    \end{bmatrix},
\end{gather}
In the absence of $\balpha$, the shifted stress $\beeta$ is
\begin{gather}
    \beeta=\begin{bmatrix}
        \dfrac{2}{3}\sigma & 0                   & 0                   \\
        0                  & -\dfrac{1}{3}\sigma & 0                   \\
        0                  & 0                   & -\dfrac{1}{3}\sigma
    \end{bmatrix},
\end{gather}
thus, its norm can be computed as
\begin{gather}
    \norm{\beeta}=\sqrt{\left(\dfrac{2}{3}\right)^2+\left(\dfrac{1}{3}\right)^2+\left(\dfrac{1}{3}\right)^2}\sigma=\sqrt{\dfrac{2}{3}}\sigma.
\end{gather}
Inserting it back, the yield function effectively defines $\sigma=\sigma^y$.

There are many equivalent alternative forms of the von Mises yield function.
For example, one can also write it as
\begin{gather}
    f=\sqrt{\dfrac{3}{2}}\norm{\beeta}-\sigma^y=\sqrt{3J_2}-\sigma^y,\qquad
    f=3J_2-\left(\sigma^y\right)^2,
\end{gather}
given that $J_2=\dfrac{1}{2}\beeta:\beeta$.
One may find all those forms (and maybe more) in the literature.
In principle, one can choose any of those forms if the yield function is the only factor considered.
However, accounting for the associated numerical implementation, which involves computing the derivative of $f$, some forms are more convenient than the others.
\subsubsection{Flow Rule}
Assuming an associative rule, the flow rule is
\begin{gather}\label{eq:j2_flow}
    \dot{\bvarepsilon^p}=\gamma\pdfrac{f}{\bsigma}=\gamma\pdfrac{\norm{\beeta}}{\bsigma}=\gamma\dfrac{1}{2}\dfrac{2\mathbb{I}^\text{dev}:\beeta}{\norm{\beeta}}=\gamma\dfrac{\beeta}{\norm{\beeta}}=\gamma\bn.
\end{gather}
All analytical formulations below are based on tensor notation.
Expressed explicitly in tensor components, the above relation becomes
\begin{gather}
    \underbrace{\begin{bmatrix}
            \dot{\varepsilon^p}_{11} & \dot{\varepsilon^p}_{12} & \dot{\varepsilon^p}_{31} \\
            \dot{\varepsilon^p}_{12} & \dot{\varepsilon^p}_{22} & \dot{\varepsilon^p}_{23} \\
            \dot{\varepsilon^p}_{31} & \dot{\varepsilon^p}_{23} & \dot{\varepsilon^p}_{33}
        \end{bmatrix}}_{\dot{\bvarepsilon^p}}
    =
    \gamma
    \underbrace{\begin{bmatrix}
            n_{11} & n_{12} & n_{31} \\
            n_{12} & n_{22} & n_{23} \\
            n_{31} & n_{23} & n_{33}
        \end{bmatrix}}_{\bn},\qquad\text{in tensor components.}
\end{gather}
We define the scaling vector $\mb{c}=\begin{bmatrix}
        1 & 1 & 1 & 2 & 2 & 2
    \end{bmatrix}^\mT$ and let $\circ$ be the Hadamard (element--wise) product operator, then the above expression is equivalent to
\begin{gather}
    \underbrace{\begin{bmatrix}
            \dot{\varepsilon^p}_{11}  \\
            \dot{\varepsilon^p}_{22}  \\
            \dot{\varepsilon^p}_{33}  \\
            2\dot{\varepsilon^p}_{12} \\
            2\dot{\varepsilon^p}_{23} \\
            2\dot{\varepsilon^p}_{31}
        \end{bmatrix}}_{\dot{\bvarepsilon^p}}
    =
    \gamma\begin{bmatrix}
        1 &   &   &   &   &   \\
          & 1 &   &   &   &   \\
          &   & 1 &   &   &   \\
          &   &   & 2 &   &   \\
          &   &   &   & 2 &   \\
          &   &   &   &   & 2
    \end{bmatrix}
    \underbrace{
        \begin{bmatrix}
            n_{11} \\
            n_{22} \\
            n_{33} \\
            n_{12} \\
            n_{23} \\
            n_{31}
        \end{bmatrix}}_{\bn}=\gamma~\mb{c}\circ\bn,\qquad\text{in the Voigt notation.}
\end{gather}
This agrees with \eqsref{eq:norm_derivative2}.
This difference has practical implications.
Because both strain and stress tensors are symmetric, it is often more efficient in implementation to use vectors and matrices to represent second--order and fourth--order tensors.
In other words, tensor quantities are expressed in compressed vector/matrix form.
For a given expression, one should always ask what operation it implies and how it will eventually be computed when the operands are expressed in different forms.
For example, if $\dot{\bvarepsilon^p}$ is used in an intermediate computation, it can simply be replaced by $\gamma\bn$ --- the mathematics remains correct as long as we use consistent tensor notation.
However, if $\bvarepsilon^p$ needs to be stored in Voigt notation, the shear components should be scaled carefully by a factor of $2$ when committing the new $\bvarepsilon^p$.

It can be observed that $\dot{\bvarepsilon^p}$ has magnitude $\gamma$, while $\bn$ is a unit tensor in the $\mathbb{R}^3\times\mathbb{R}^3$ hyperspace. Thus, $\bn$ serves as a direction indicator, and $\bvarepsilon^p$ evolves towards $\bn$ by the amount $\gamma$. Since $\bn$ is deviatoric, $\dot{\bvarepsilon^p}$ is also deviatoric (and so is $\bvarepsilon^p$).
Furthermore, since $\bn$ is a unit tensor,
\begin{gather}
    \norm{\dot{\bvarepsilon^p}}=\gamma.
\end{gather}
\subsubsection{Hardening Law}
For the internal variable $q$, the hardening law uses the accumulated magnitude of $\bvarepsilon^p$.
\begin{gather}
    \dot{q}=\sqrt{\dfrac{2}{3}}\norm{\dot{\bvarepsilon^p}}=\sqrt{\dfrac{2}{3}}\gamma.
\end{gather}
Similar to the stress counterpart, here the constant $\sqrt{\dfrac{2}{3}}$ needs to be introduced to ensure a consistent and correct uniaxial response.
To see this, consider the uniaxial loading case again, to have a unit plastic strain increment along the designated axis (say, the $x$-axis), also noting that $\dot{\bvarepsilon^p}$ is \textbf{deviatoric}, the full plastic strain increment tensor must be
\begin{gather}
    \dot{\bvarepsilon^p}=\dfrac{1}{2}\begin{bmatrix}
        2 & -1 & -1 & 0 & 0 & 0
    \end{bmatrix}^\mT.
\end{gather}
Clearly, to have a unit first component, the other two normal components must be $-\dfrac{1}{2}$ each.
This will lead to $\norm{\dot{\bvarepsilon^p}}=\sqrt{\dfrac{3}{2}}$.
If we still accept the definition $\dot{q}=\norm{\dot{\bvarepsilon^p}}$, then the same unit plastic strain increment that yields $\dot{q}=1$ in the uniaxial formulation would now give $\dot{q}=\sqrt{\dfrac{3}{2}}$ in the 3D formulation.
As a result, the same hardening function would produce different responses in two cases.
To have the identical $\dot{q}$, we must scale down $\norm{\dot{\bvarepsilon^p}}$ by $\sqrt{\dfrac{2}{3}}$, thus the above expression is obtained.

For isotropic hardening, $\sigma^y$ is defined as a general function of $q$,
\begin{gather}
    \sigma^y=\sigma^y\left(q\right).
\end{gather}

For kinematic hardening, the evolution law of back stress $\balpha$ is defined to be
\begin{gather}
    \dot{\balpha}=\sqrt{\dfrac{2}{3}}\dot{H}\bn.
\end{gather}
Here, $H=H\left(q\right)$ is now a scalar--valued function of $q$ that controls the development of $\norm{\balpha}$. The tensor $\balpha$ always evolves towards $\bn$ by the amount $\dot{H}=H\left(q_{n+1}\right)-H\left(q_n\right)$, which characterises the increment of $H$. The factor $\sqrt{\dfrac{2}{3}}$ is introduced for a consistent response, as stated earlier. Since $\dot{\balpha}$ and $\bn$ are coaxial, $\balpha$ stays deviatoric but may not remain coaxial with all $\bn$ throughout the loading process.
\subsection{Formulation}
A summary of the von Mises model is listed below.
\begin{table}[ht]
    \centering
    \begin{tabular}{rl}
        \toprule
        Constitutive Law & $\bsigma=\mb{D}:\left(\bvarepsilon-\bvarepsilon^p\right)$ \\
        Yield Function   & $f=\norm{\beeta}-\sqrt{\dfrac{2}{3}}\sigma^y$             \\
        Flow Rule        & $\dot{\bvarepsilon^p}=\gamma\bn$                          \\
        Hardening Law    & $\dot{q}=\sqrt{\dfrac{2}{3}}\gamma$                       \\
                         & $\dot{\balpha}=\sqrt{\dfrac{2}{3}}\dot{H}\bn$             \\\bottomrule
    \end{tabular}
\end{table}
\subsubsection{Elastic Loading/Unloading}
The trial stress can be computed as
\begin{gather}\label{eq:j2_tsigma}
    \bsigma^\text{trial}=\mb{D}:\left(\bvarepsilon_{n+1}-\bvarepsilon^p_n\right)=\bsigma_n+\mb{D}:\left(\bvarepsilon_{n+1}-\bvarepsilon_n\right).
\end{gather}
In matrix representation, it is
\begin{gather}
    \underbrace{
        \begin{bmatrix}
            \sigma^\text{trial}_{11} \\
            \sigma^\text{trial}_{22} \\
            \sigma^\text{trial}_{33} \\
            \sigma^\text{trial}_{12} \\
            \sigma^\text{trial}_{23} \\
            \sigma^\text{trial}_{31}
        \end{bmatrix}}_{\bsigma^\text{trial}}
    =
    \underbrace{\begin{bmatrix}
            \lambda+2G & \lambda    & \lambda    & \cdot & \cdot & \cdot \\
            \lambda    & \lambda+2G & \lambda    & \cdot & \cdot & \cdot \\
            \lambda    & \lambda    & \lambda+2G & \cdot & \cdot & \cdot \\
            \cdot      & \cdot      & \cdot      & 2G    & \cdot & \cdot \\
            \cdot      & \cdot      & \cdot      & \cdot & 2G    & \cdot \\
            \cdot      & \cdot      & \cdot      & \cdot & \cdot & 2G
        \end{bmatrix}}_{\mb{D}}
    \underbrace{\left(\begin{bmatrix}
            \varepsilon_{11,n+1} \\
            \varepsilon_{22,n+1} \\
            \varepsilon_{33,n+1} \\
            \varepsilon_{12,n+1} \\
            \varepsilon_{23,n+1} \\
            \varepsilon_{31,n+1}
        \end{bmatrix}-\begin{bmatrix}
            \varepsilon^p_{11,n} \\
            \varepsilon^p_{22,n} \\
            \varepsilon^p_{33,n} \\
            \varepsilon^p_{12,n} \\
            \varepsilon^p_{23,n} \\
            \varepsilon^p_{31,n}
        \end{bmatrix}
        \right).}_{\bvarepsilon_{n+1}-\bvarepsilon^p_n}
\end{gather}
Using engineering strains, the above must be adjusted to
\begin{gather}
    \underbrace{
        \begin{bmatrix}
            \sigma^\text{trial}_{11} \\
            \sigma^\text{trial}_{22} \\
            \sigma^\text{trial}_{33} \\
            \sigma^\text{trial}_{12} \\
            \sigma^\text{trial}_{23} \\
            \sigma^\text{trial}_{31}
        \end{bmatrix}}_{\bsigma^\text{trial}}
    =
    \underbrace{\begin{bmatrix}
            \lambda+2G & \lambda    & \lambda    & \cdot & \cdot & \cdot \\
            \lambda    & \lambda+2G & \lambda    & \cdot & \cdot & \cdot \\
            \lambda    & \lambda    & \lambda+2G & \cdot & \cdot & \cdot \\
            \cdot      & \cdot      & \cdot      & G     & \cdot & \cdot \\
            \cdot      & \cdot      & \cdot      & \cdot & G     & \cdot \\
            \cdot      & \cdot      & \cdot      & \cdot & \cdot & G
        \end{bmatrix}}_{\mb{D}\text{ or conventionally }\mb{E}}
    \underbrace{\left(\begin{bmatrix}
            \varepsilon_{11,n+1} \\
            \varepsilon_{22,n+1} \\
            \varepsilon_{33,n+1} \\
            \gamma_{12,n+1}      \\
            \gamma_{23,n+1}      \\
            \gamma_{31,n+1}
        \end{bmatrix}-\begin{bmatrix}
            \varepsilon^p_{11,n} \\
            \varepsilon^p_{22,n} \\
            \varepsilon^p_{33,n} \\
            \gamma^p_{12,n}      \\
            \gamma^p_{23,n}      \\
            \gamma^p_{31,n}
        \end{bmatrix}
        \right).}_{\bvarepsilon_{n+1}-\bvarepsilon^p_n}
\end{gather}

Then $\beeta^\text{trial}=\dev{\bsigma^\text{trial}}-\balpha_n$, which gives the trial yield function
\begin{gather}\label{eq:j2_tf}
    f^\text{trial}=\norm{\beeta^\text{trial}}-\sqrt{\dfrac{2}{3}}\sigma^y_n
\end{gather}
with $\sigma^y_n=\sigma^y\left(q_n\right)$ evaluated with current $q_n$.
The above expression simply assumes the current loading increment $\Delta\bvarepsilon=\bvarepsilon_{n+1}-\bvarepsilon_n$ is purely elastic, thus all plasticity related variables remain unchanged.
The yield function evaluated in this way gives an upper bound of the actual yield function.
If this value is negative, the actual state must be elastic, otherwise, plasticity evolution must occur.
\subsubsection{Plastic Evolution}
By default, we present the formulation with the implicit Euler method.

If plasticity evolution occurs, that is, $f^\text{trial}>0$, $\gamma$ must be non-zero and thus needs to be solved.
\begin{gather}
    \beeta_{n+1}=\dev{\bsigma_{n+1}}-\balpha_{n+1}.
\end{gather}
Accounting for
\begin{gather}
    \bsigma_{n+1}=\bsigma^\text{trial}-\mb{D}:\dot{\bvarepsilon^p}=\bsigma^\text{trial}-\gamma2G\bn,\\
    \balpha_{n+1}=\balpha_n+\sqrt{\dfrac{2}{3}}\dot{H}\bn,
\end{gather}
one can further derive
\begin{gather}
    \beeta_{n+1}=\dev{\bsigma^\text{trial}-\gamma2G\bn}-\balpha_n-\sqrt{\dfrac{2}{3}}\dot{H}\bn.
\end{gather}
Since $\bn$ is already deviatoric, meaning $\dev{\bn}=\bn$, then,
\begin{gather}
    \begin{split}
        \beeta_{n+1} & =\dev{\bsigma^\text{trial}}-\gamma2G\bn-\balpha_n-\sqrt{\dfrac{2}{3}}\dot{H}\bn \\
                     & =\beeta^\text{trial}-\left(2G\gamma+\sqrt{\dfrac{2}{3}}\dot{H}\right)\bn.
    \end{split}
\end{gather}
Noting that, by definition, $\bn=\dfrac{\beeta_{n+1}}{\norm{\beeta_{n+1}}}$, thus,
\begin{gather}
    \underbrace{\left(1+\dfrac{2G\gamma+\sqrt{\dfrac{2}{3}}\dot{H}}{\norm{\beeta_{n+1}}}\right)}_{\text{scalar}}\beeta_{n+1}=\beeta^\text{trial}.
\end{gather}
The above expression implies that $\beeta_{n+1}$ and $\beeta^\text{trial}$ are \textbf{coaxial}. Thus, $\beeta_{n+1}$ always points in the direction of $\beeta^\text{trial}$, which remains constant within each time step.
Only its magnitude is adjusted during local iterations.
Thus, $\bn=\dfrac{\beeta_{n+1}}{\norm{\beeta_{n+1}}}=\dfrac{\beeta^\text{trial}}{\norm{\beeta^\text{trial}}}$ is also constant during the local iteration.
This is a convenient property, although it may not exist in more complex models.
However, when it is available, it can be exploited to simplify the formulation.
One can further derive
\begin{gather}
    \norm{\beeta_{n+1}}+2G\gamma+\sqrt{\dfrac{2}{3}}\dot{H}=\norm{\beeta^\text{trial}}.
\end{gather}

The yield function evaluated at the new state thus reads
\begin{gather}\label{eq:j2_f}
    f=\norm{\beeta^\text{trial}}-\left(2G\gamma+\sqrt{\dfrac{2}{3}}\left(H_{n+1}-H_n\right)\right)-\sqrt{\dfrac{2}{3}}\sigma^y_{n+1}.
\end{gather}

The Jacobian reads
\begin{gather}
    \pdfrac{f}{\gamma}=-2G-\sqrt{\dfrac{2}{3}}\ddfrac{H_{n+1}}{q_{n+1}}\ddfrac{q_{n+1}}{\gamma}-\sqrt{\dfrac{2}{3}}\ddfrac{\sigma^y_{n+1}}{q_{n+1}}\ddfrac{q_{n+1}}{\gamma}.
\end{gather}
Since $q_{n+1}=q_n+\sqrt{\dfrac{2}{3}}\gamma$, $\ddfrac{q_{n+1}}{\gamma}=\sqrt{\dfrac{2}{3}}$. Hence,
\begin{gather}\label{eq:j2_df}
    \pdfrac{f}{\gamma}=-2G-\dfrac{2}{3}\ddfrac{H_{n+1}}{q_{n+1}}-\dfrac{2}{3}\ddfrac{\sigma^y_{n+1}}{q_{n+1}}.
\end{gather}
\subsubsection{Consistent Tangent Stiffness}
From $\bsigma_{n+1}=\bsigma^\text{trial}-\gamma2G\bn$, as $\bn=\dfrac{\beeta}{\norm{\beeta}}=\dfrac{\beeta^\text{trial}}{\norm{\beeta^\text{trial}}}$, the consistent tangent stiffness can be computed via the chain rule as
\begin{gather}
    \pdfrac{\bsigma_{n+1}}{\bvarepsilon_{n+1}}=\pdfrac{\bsigma_{n+1}^\text{trial}}{\bvarepsilon_{n+1}}-2G\pdfrac{\left(\gamma\bn\right)}{\bvarepsilon_{n+1}}=\mb{D}-2G\left(\bn\otimes\pdfrac{\gamma}{\bvarepsilon_{n+1}}+\gamma\pdfrac{\bn}{\bvarepsilon_{n+1}}\right).
\end{gather}
In which, according to \eqsref{eq:unit_derivative},
\begin{gather}
    \begin{split}
        \pdfrac{\bn}{\bvarepsilon_{n+1}} & =\dfrac{1}{\norm{\beeta^\text{trial}}}\left(\mathbb{I}-\bn\otimes\bn\right):\pdfrac{\mb{\beeta}^\text{trial}}{\bvarepsilon_{n+1}} \\
                                         & =\dfrac{2G}{\norm{\beeta^\text{trial}}}\left(\mathbb{I}-\bn\otimes\bn\right):\mathbb{I}^\text{dev}                                \\
                                         & =\dfrac{2G}{\norm{\beeta^\text{trial}}}\left(\mathbb{I}^\text{dev}-\bn\otimes\bn\right).
    \end{split}
\end{gather}
Here, $\bn$ is expressed as $\bn=\dfrac{\beeta^\text{trial}}{\norm{\beeta^\text{trial}}}$ because of the coaxiality between $\beeta_{n+1}$ and $\beeta^\text{trial}$.
This simplifies the computation.

From the converged local residual (yield function),
\begin{gather}
    \begin{split}
        \pdfrac{\gamma}{\bvarepsilon_{n+1}} & =-\left(\pdfrac{f}{\gamma}\right)^{-1}\pdfrac{f}{\bvarepsilon_{n+1}}=-\left(\pdfrac{f}{\gamma}\right)^{-1}2G\bn.
    \end{split}
\end{gather}

Thus the final expression for consistent tangent stiffness can be assembled as
\begin{gather}\label{eq:j2_stiffness}
    \begin{split}
        \pdfrac{\bsigma_{n+1}}{\bvarepsilon_{n+1}} & =\mb{D}-2G\left(-2G\left(\pdfrac{f}{\gamma}\right)^{-1}\bn\otimes\bn+\gamma\dfrac{2G}{\norm{\beeta^\text{trial}}}\left(\mathbb{I}^\text{dev}-\bn\otimes\bn\right)\right)                                                                               \\
                                                   & =\mb{D}+4G^2\left(\pdfrac{f}{\gamma}\right)^{-1}\bn\otimes\bn+\dfrac{4G^2\gamma}{\norm{\beeta^\text{trial}}}\left(\bn\otimes\bn-\mathbb{I}^\text{dev}\right)                                                                                           \\
                                                   & =\mb{D}+4G^2\left(\left(\pdfrac{f}{\gamma}\right)^{-1}+\dfrac{\gamma}{\norm{\beeta^\text{trial}}}\right)\bn\otimes\bn-\dfrac{4G^2\gamma}{\norm{\beeta^\text{trial}}}\mathbb{I}^\text{dev}                                                              \\
                                                   & =\mb{D}+4G^2\left(\dfrac{\gamma}{\norm{\beeta^\text{trial}}}-\dfrac{1}{2G+\dfrac{2}{3}\ddfrac{H_{n+1}}{q_{n+1}}+\dfrac{2}{3}\ddfrac{\sigma^y_{n+1}}{q_{n+1}}}\right)\bn\otimes\bn-\dfrac{4G^2\gamma}{\norm{\beeta^\text{trial}}}\mathbb{I}^\text{dev}.
    \end{split}
\end{gather}
It is a symmetric tensor/matrix.
Note that $\mathbb{I}^\text{dev}$ takes the form presented in \eqsref{eq:dev_tensor_se}.
Readers are strongly encouraged to derive it in both tensor notation and compressed matrix representation as practice.
Both should lead to identical results.

Since the local iteration is a scalar function, the closed--form consistent tangent stiffness is relatively easy to compute. In more complex models, closed forms are not always so simple.

As a general framework, the above formulation does not require explicit forms of $H\left(q\right)$ and $\sigma^y\left(q\right)$. Thus, various types of scalar--valued functions can be adopted.
\subsection{Implementation}
The state determination algorithm for this general model with the von Mises criterion is given in \algoref{algo:j2_model}.
\begin{breakablealgorithm}
    \caption{state determination of general von Mises model}\label{algo:j2_model}
    \begin{algorithmic}
        \State \textbf{Parameter}: $\lambda$, $G$
        \State \textbf{Input}: $\bvarepsilon_{n+1}$, $\bvarepsilon_n$, $\bvarepsilon^p_n$, $\bsigma_n$, $\balpha_n$, $q_n$
        \State \textbf{Output}: $\mb{D}_{n+1}$, $\bvarepsilon^p_{n+1}$, $\bsigma_{n+1}$, $\balpha_{n+1}$, $q_{n+1}$
        \State compute $\bsigma^\text{trial}$, $\beeta^\text{trial}$, $\bn$ and $f^\text{trial}$\Comment{\eqsref{eq:j2_tsigma} and \eqsref{eq:j2_tf}}
        \If {$f^\text{trial}\geqslant0$}
        \State $\gamma=0$
        \While{true}
        \State compute $f$ and $\pdfrac{f}{\gamma}$\Comment{\eqsref{eq:j2_f} and \eqsref{eq:j2_df}}
        \State update $\Delta{}H=H\left(q_{n+1}\right)-H\left(q_n\right)$
        \State $\Delta\gamma=\left(\pdfrac{f}{\gamma}\right)^{-1}f$
        \If {$\abs{\Delta\gamma}<\text{tolerance}$}
        \State break
        \EndIf
        \State $\gamma\leftarrow\gamma-\Delta\gamma$
        \State $q_{n+1}=q_n+\sqrt{\dfrac{2}{3}}\gamma$
        \EndWhile
        \State $\bsigma_{n+1}=\bsigma^\text{trial}-\gamma2G\bn$
        \State $\bvarepsilon^p_{n+1}=\bvarepsilon^p_n+\gamma\bn$
        \State $\balpha_{n+1}=\balpha_n+\sqrt{\dfrac{2}{3}}\Delta{}H\bn$
        \State $q_{n+1}=q_n+\sqrt{\dfrac{2}{3}}\gamma$
        \State compute $\mb{D}_{n+1}$\Comment{\eqsref{eq:j2_stiffness}}
        \Else
        \State $\bsigma_{n+1}=\bsigma^\text{trial}$
        \State $\bvarepsilon^p_{n+1}=\bvarepsilon^p_n$
        \State $\balpha_{n+1}=\balpha_n$
        \State $q_{n+1}=q_n$
        \State $\mb{D}_{n+1}=\mb{D}$
        \EndIf
    \end{algorithmic}
\end{breakablealgorithm}
Note that the algorithm does not implement $H\left(q\right)$ and $\sigma^y\left(q\right)$. It assumes that these two functions are defined elsewhere and can provide derivatives.

Depending on the stress update used in \eqsref{eq:j2_tsigma}, the history of $\bvarepsilon^p$ is \textbf{not} compulsory in this model if the latter expression is used instead of the former.
\subsection{Closing Remarks}
As the first 3D material model introduced, the von Mises framework allows beginners to become familiar with multiaxial constitutive modelling with a relatively smooth learning curve. The formulation is expressed in tensor notation. Readers are strongly encouraged to derive the formulation independently from the three governing equations in both tensor and compressed matrix notations. It is good practice to get every small detail correct.
\begin{cppcode}
int NonlinearJ2::update_trial_status(const vec& t_strain) {
    incre_strain = (trial_strain = t_strain) - current_strain;

    if(norm(incre_strain) <= datum::eps) return SUANPAN_SUCCESS;

    trial_stress = current_stress + (trial_stiffness = initial_stiffness) * incre_strain;

    trial_history = current_history;
    auto& plastic_strain = trial_history(0);
    vec back_stress(&trial_history(1), 6, false, true);

    const vec rel_stress = tensor::dev(trial_stress) - back_stress;
    const auto norm_rel_stress = tensor::stress::norm(rel_stress);

    double k, dk;

    auto update_isotropic_hardening = [&] {
        k = compute_k(plastic_strain);
        dk = 0.;
        k < 0. ? k = 0. : dk = compute_dk(plastic_strain);
    };

    update_isotropic_hardening();

    auto yield_func = norm_rel_stress - root_two_third * k;

    if(yield_func <= 0.) return SUANPAN_SUCCESS;

    const auto current_h = compute_h(plastic_strain);
    auto gamma = 0., incre_h = 0.;
    double denom;
    auto counter = 0u;
    auto ref_error = 1.;
    while(true) {
        if(max_iteration == ++counter) {
            suanpan_error("Cannot converge within {} iterations.\n", max_iteration);
            return SUANPAN_FAIL;
        }

        denom = double_shear + two_third * (dk + compute_dh(plastic_strain));
        const auto incre_gamma = yield_func / denom;
        const auto error = std::fabs(incre_gamma);
        if(1u == counter) ref_error = error;
        suanpan_debug("Local iteration error: {:.5E}.\n", error);
        if(error < tolerance * ref_error || ((error < tolerance || std::fabs(yield_func) < tolerance) && counter > 5u)) break;
        incre_h = compute_h(plastic_strain = current_history(0) + root_two_third * (gamma += incre_gamma)) - current_h;
        update_isotropic_hardening();
        yield_func = norm_rel_stress - double_shear * gamma - root_two_third * (k + incre_h);
    }

    back_stress += root_two_third * incre_h / norm_rel_stress * rel_stress;

    auto t_factor = double_shear * gamma / norm_rel_stress;
    trial_stress -= t_factor * rel_stress;

    t_factor *= double_shear;
    trial_stiffness += (t_factor - square_double_shear / denom) / norm_rel_stress / norm_rel_stress * rel_stress * rel_stress.t() - t_factor * unit_dev_tensor;

    return SUANPAN_SUCCESS;
}
\end{cppcode}

Finally, in \eqsref{eq:j2_flow}, we have discussed that the same expression
\begin{gather}
    \dot{\bvarepsilon^p}=\gamma\bn
\end{gather}
may have different explicit forms, depending on how those quantities are represented.

This $\dot{\bvarepsilon^p}$ is further used in
\begin{gather}\label{eq:j2_two_form}
    \bs^\text{trial}-\bs_{n+1}=2G\dot{\bvarepsilon^p}.
\end{gather}
The above expression assumes the tensor notation.
If the matrix notation is used, it can be explicitly written as
\begin{gather}
    \begin{bmatrix}
        s^\text{trial}_{11}-s_{11,n+1} \\
        s^\text{trial}_{22}-s_{22,n+1} \\
        s^\text{trial}_{33}-s_{33,n+1} \\
        s^\text{trial}_{12}-s_{12,n+1} \\
        s^\text{trial}_{23}-s_{23,n+1} \\
        s^\text{trial}_{31}-s_{31,n+1}
    \end{bmatrix}=\gamma\begin{bmatrix}
        2G &    &    &   &   &   \\
           & 2G &    &   &   &   \\
           &    & 2G &   &   &   \\
           &    &    & G &   &   \\
           &    &    &   & G &   \\
           &    &    &   &   & G
    \end{bmatrix}\diag{\mb{c}}\bn.
\end{gather}
One immediately realises that
\begin{gather}
    \diag{\begin{matrix}
            2G \\2G\\2G\\G\\G\\G
        \end{matrix}}\circ\diag{\mb{c}}\bn=\diag{\begin{matrix}
            2G \\2G\\2G\\2G\\2G\\2G
        \end{matrix}}\bn=2G\bn,
\end{gather}
thus, \eqsref{eq:j2_two_form} becomes
\begin{gather}
    \bs^\text{trial}-\bs_{n+1}=\gamma2G\bn.
\end{gather}
No matter which notation is used, the same expression holds.
This is based on the fact that $\dot{\bvarepsilon^p}$ is used as an intermediate variable.
The intermediate scaling matrices cancel out and eventually the same expression is obtained.

This is a useful attribute, although it can be confusing for beginners.
